\vspace{-2.5em} Singular value decomposition of single view\\ polarized fluorescence microscopes\vspace{-1em}

### 2018-05-22 notes/2018-05-14-single-view-continuous-svd/report/report.tex
Circular harmonic coefficients.

\noindent We will make use of the real circular harmonic functions
\begin{align}
  z_n(\theta) =
  \begin{cases}
    \frac{1}{\sqrt{\pi}}\cos(n\theta), & n > 0,\\
    \frac{1}{\sqrt{\pi}}, & n = 0,\\
    \frac{1}{\sqrt{\pi}}\sin(n\theta), & n < 0.\\
  \end{cases}
\end{align}
The real circular harmonic functions form an orthonormal basis for functions on
the circle because
\begin{align}
  \int_{\mbb{S}^1}d\mh{p}\, z_n(\mh{p})z_{n'}(\mh{p}) = \delta_{nn'}.
\end{align}
Notice that we allow $z_n(\cdot)$ to take a vector argument similar to the
spherical harmonics $y_l^m(\so{})$. This allows us to use these functions
without specifying a coordinate system. Also notice that we are using $n$ to
index the circular harmonic functions---in a previous note set we used $n$ to
index frames or views.

@@

We also showed that the point response function for polarized detection is given
by (rewritten in terms of the circular harmonics)
\begin{align}
  h_{\text{det}}(\ro{}, \so{}; \mh{p}) = \sum_{l=0}^{\infty}\sum_{m=-l}^l\sum_{n=-\infty}^{\infty} h_{\text{det},l,n}^m(\ro{})y_l^m(\so{})z_n(\mh{p}_d), 
\end{align}
where
\begin{align}
  h_{\text{det},0,0}^0(\ro{}) &= a^2(r_o) + 2b^2(r_0),\\
  h_{\text{det},0,\pm 2}^0(\ro{}) &= 2\sqrt{\pi} b^2(r_0)z_{\pm 2}(\ro{}),\\
  h_{\text{det},2,0}^0(\ro{}) &= \frac{1}{\sqrt{5}}\left[-a^2(r_o) + 4b^2(r_0)\right],\\
  h_{\text{det},2,\pm 2}^0(\ro{}) &= 4\sqrt{\frac{\pi}{5}}b^2(r_0)z_{\pm 2}(\ro{}),\\
  h_{\text{det},2,2}^2(\ro{}) &= -h_{\text{det},2,-2}^{-2}(\ro{}) = \sqrt{\frac{3}{5}}a^2(r_o).
\end{align}
The complete kernel for unpolarized illumination and polarized detection is given by
\begin{align}
  h(\ro{}, \so{}; \mh{p}_d) = h_{\text{exc}}(\so{})h_{\text{det}}(\ro{}, \so{}; \mh{p}_d) = \sum_{l=0}^{\infty}\sum_{m=-l}^l\sum_{n=-\infty}^{\infty} h_{l,n}^m(\ro{})y_l^m(\so{})z_n(\mh{p}_d), 
\end{align}
where
\begin{align}
  h_{0,0}^0(\ro{}) &= h_{\text{exc},0,0}^0 h_{\text{det},0,0}^0(\ro{}) + h_{\text{exc},2,0}^0 h_{\text{det},2,0}^0(\ro{}),\\
  h_{0,\pm 2}^0(\ro{}) &= h_{\text{exc},0,0}^0 h_{\text{det},0,\pm 2}^0(\ro{}) + h_{\text{exc},2,0}^0 h_{\text{det},2,\pm 2}^0(\ro{}),\\
  h_{2,0}^0(\ro{}) &= h_{\text{exc},0,0}^0 h_{\text{det},0,0}^0(\ro{}) + h_{\text{exc},2,0}^0 h_{\text{det},2,0}^0(\ro{}) + \frac{2\sqrt{5}}{7}h_{\text{exc},2,0}^0h_{\text{det},2,0}^0(\ro{}),\\
  h_{2,\pm 2}^0(\ro{}) &= h_{\text{exc},0,0}^0 h_{\text{det},0,\pm 2}^0(\ro{}) + h_{\text{exc},2,0}^0 h_{\text{det},2,\pm 2}^0(\ro{}) + \frac{2\sqrt{5}}{7}h_{\text{exc},2,0}^0h_{\text{det},2,\pm 2}^0(\ro{}),\\
  h_{2,\pm 2}^{\pm 2}(\ro{}) &= h_{\text{det},2,\pm 2}^{\pm 2}(\ro{})\left[h_{\text{exc},0,0}^0 - \frac{2\sqrt{5}}{7}h_{\text{exc},2,0}^0\right],\\
  h_{4,0}^0(\ro{}) &= \frac{6}{7}h_{\text{exc},2,0}^0h_{\text{det},2,0}^0(\ro{}),\\
  h_{4,\pm 2}^0(\ro{}) &= \frac{6}{7}h_{\text{exc},2,0}^0h_{\text{det},2,\pm 2}^0(\ro{}),\\
  h_{4,\pm 2}^{\pm 2}(\ro{}) &= \frac{\sqrt{15}}{7}h_{\text{exc},2,0}^0h_{\text{det},2,\pm 2}^{\pm 2}(\ro{}),
\end{align}
% hh0
% [h000d*h000e + h200d*h200e, 0, 0, h000d*h200e + h000e*h200d + 2*sqrt(5)*h200d*h200e/7, 0, 0, 0, 0, 0, 0, 6*h200d*h200e/7, 0, 0, 0, 0]

@@

\end{document}

>>>>

\noindent We will make use of the real circular harmonic functions
\begin{align}
  z_n(\theta) =
  \begin{cases}
    \frac{1}{\sqrt{\pi}}\cos(n\theta), & n > 0,\\
    \frac{1}{\sqrt{2\pi}}, & n = 0,\\
    \frac{1}{\sqrt{\pi}}\sin(n\theta), & n < 0.\\
  \end{cases}
\end{align}
The real circular harmonic functions form an orthonormal basis for functions on
the circle because
\begin{align}
  \int_{\mbb{S}^1}d\mh{p}\, z_n(\mh{p})z_{n'}(\mh{p}) = \delta_{nn'}.
\end{align}
Notice that we allow $z_n(\cdot)$ to take a vector argument similar to the
spherical harmonics $y_l^m(\so{})$. This allows us to use these functions
without specifying a coordinate system. Also notice that we are using $n$ to
index the circular harmonic functions---in a previous note set we used $n$ to
index frames or views.

@@

We also showed that the point response function for polarized detection is given
by (rewritten in terms of the circular harmonics)
\begin{align}
  h_{\text{det}}(\ro{}, \so{}; \mh{p}) = \sum_{l=0}^{\infty}\sum_{m=-l}^l\sum_{n=-\infty}^{\infty} h_{\text{det},l,n}^m(\ro{})y_l^m(\so{})z_n(\mh{p}_d), 
\end{align}
where
\begin{align}
  h_{\text{det},0,0}^0(\ro{}) &= a^2(r_o) + 2b^2(r_0),\\
  h_{\text{det},0,\pm 2}^0(\ro{}) &= 2\sqrt{\pi} b^2(r_0)z_{\pm 2}(\ro{}),\\
  h_{\text{det},2,0}^0(\ro{}) &= \frac{1}{\sqrt{5}}\left[-a^2(r_o) + 4b^2(r_0)\right],\\
  h_{\text{det},2,\pm 2}^0(\ro{}) &= 4\sqrt{\frac{\pi}{5}}b^2(r_0)z_{\pm 2}(\ro{}),\\
  h_{\text{det},2,2}^2(\ro{}) &= -h_{\text{det},2,-2}^{-2}(\ro{}) = \sqrt{\frac{3}{5}}a^2(r_o).
\end{align}
The complete kernel for unpolarized illumination and polarized detection is
given by the product of the excitation and detection kernels. In the previous
case we were able to calculate this product in terms of the circular and
spherical harmonics by hand, but this task is difficult in the general case. To
simplify this task we precompute the triple integrals of the circular and spherical
harmonics---see Appendix A for details. 
\begin{align}
  h(\ro{}, \so{}; \mh{p}_d) = h_{\text{exc}}(\so{})h_{\text{det}}(\ro{}, \so{}; \mh{p}_d) = \sum_{l=0}^{\infty}\sum_{m=-l}^l\sum_{n=-\infty}^{\infty} h_{l,n}^m(\ro{})y_l^m(\so{})z_n(\mh{p}_d), 
\end{align}
where
\begin{align}
  h_{0,0}^0(\ro{}) &= h_{\text{exc},0,0}^0 h_{\text{det},0,0}^0(\ro{}) + h_{\text{exc},2,0}^0 h_{\text{det},2,0}^0(\ro{}),\\
  h_{0,\pm 2}^0(\ro{}) &= h_{\text{exc},0,0}^0 h_{\text{det},0,\pm 2}^0(\ro{}) + h_{\text{exc},2,0}^0 h_{\text{det},2,\pm 2}^0(\ro{}),\\
  h_{2,0}^0(\ro{}) &= h_{\text{exc},0,0}^0 h_{\text{det},0,0}^0(\ro{}) + h_{\text{exc},2,0}^0 h_{\text{det},2,0}^0(\ro{}) + \frac{2\sqrt{5}}{7}h_{\text{exc},2,0}^0h_{\text{det},2,0}^0(\ro{}),\\
  h_{2,\pm 2}^0(\ro{}) &= h_{\text{exc},0,0}^0 h_{\text{det},0,\pm 2}^0(\ro{}) + h_{\text{exc},2,0}^0 h_{\text{det},2,\pm 2}^0(\ro{}) + \frac{2\sqrt{5}}{7}h_{\text{exc},2,0}^0h_{\text{det},2,\pm 2}^0(\ro{}),\\
  h_{2,\pm 2}^{\pm 2}(\ro{}) &= h_{\text{det},2,\pm 2}^{\pm 2}(\ro{})\left[h_{\text{exc},0,0}^0 - \frac{2\sqrt{5}}{7}h_{\text{exc},2,0}^0\right],\\
  h_{4,0}^0(\ro{}) &= \frac{6}{7}h_{\text{exc},2,0}^0h_{\text{det},2,0}^0(\ro{}),\\
  h_{4,\pm 2}^0(\ro{}) &= \frac{6}{7}h_{\text{exc},2,0}^0h_{\text{det},2,\pm 2}^0(\ro{}),\\
  h_{4,\pm 2}^{\pm 2}(\ro{}) &= \frac{\sqrt{15}}{7}h_{\text{exc},2,0}^0h_{\text{det},2,\pm 2}^{\pm 2}(\ro{}),
\end{align}
% hh0
% [h000d*h000e + h200d*h200e, 0, 0, h000d*h200e + h000e*h200d + 2*sqrt(5)*h200d*h200e/7, 0, 0, 0, 0, 0, 0, 6*h200d*h200e/7, 0, 0, 0, 0]

@@

\appendix
\section{Calculating products of circular and spherical harmonics}
To calculate the kernels of arbitrary microscope designs we need an efficient
way to calculate the products of functions that are linear combinations of
circular and spherical harmonics. We start by considering the simplest case of
multiplying two linear combinations of circular harmonics given by
\begin{align}
  f(\mh{p}) = \sum_{n=0}^\infty c_nz_n(\mh{p}), \qquad f'(\mh{p}) = \sum_{n'=0}^\infty c_{n'}'z_{n'}(\mh{p}). 
\end{align}
The product of these two functions is given by 
\begin{align}
  f''(\mh{p}) &= f(\mh{p})f'(\mh{p})=\sum_{n=0}^\infty \sum_{n'=0}^\infty c_n c_{n'}' z_n(\mh{p}) z_{n'}(\mh{p}) = \sum_{n''=0}^{\infty} c_{n''}''z_{n''}(\mh{p}), \label{eq:two}
  \end{align}
  and we would like to find a relationship between the input coefficients ($c_n$
  $c_{n'}'$) and the output coefficients $c_{n''}''$. To find this relationship
  we can rewrite the product the circular harmonics as a linear combination of
  circular harmonics
  \begin{align}
    z_n(\mh{p})z_{n'}(\mh{p}) = \sum_{j''=0}^{\infty}P_{n,n',n''}z_{j''}(\mh{p})\label{eq:product}
  \end{align}
  where $P^{n''}_{n,n'}$ are the triple integrals of the circular harmonics
  \begin{align}
    P_{n,n',n''} = \int_{\mbb{S}^1}d\mh{p}\, z_{n}(\mh{p})z_{n'}(\mh{p})z_{n''}(\mh{p}).\label{eq:triple}
  \end{align}
  Plugging Eq. \ref{eq:product} into Eq. \ref{eq:two} gives
  \begin{align}
    \sum_{n=0}^\infty \sum_{n'=0}^\infty c_n c_{n'}' \left[\sum_{j''=0}^{\infty}P^{n''}_{n,n'}z_{j''}(\mh{p})\right] = \sum_{n''=0}^{\infty} c_{n''}''z_{n''}(\mh{p}).    
  \end{align}
  Therefore, the coefficients are related by
  \begin{align}
    c_{n''}'' = \sum_{n=0}^\infty \sum_{n'=0}^\infty P_{n,n',n''} c_n c_{n'}'.
  \end{align}
  Rewriting this equation in Einstein notation (summation over matching indices
  is implied, upper indices are ``column'' indices, lower indices are ``row''
  indices) gives
  \begin{align}
    {c''}^{n''} = P^{n''}_{n,n'} c_n c_{n'}'.
  \end{align}
  We can precompute the triple integrals $P_{n,n',n''}$ using a symbolic package
  like Sympy.

  The discussion above applies to spherical harmonics as well---we only need to replace
  the triple integral of the circular harmonics $P_{n,n',n''}$ with the triple integral
  of the spherical harmonics
  \begin{align}
    G_{j,j',j''} = \int_{\mbb{S}^2}d\mh{s}\, y_{j}(\mh{s})y_{j'}(\mh{s})y_{j''}(\mh{s}).
  \end{align}
  where each of the $j$ indices are a single index over the spherical harmonics.
  We could compute the triple integrals symbolically, but these integrals can
  take several minutes. Instead, we can write the integral in terms of the Gaunt
  coefficients \cite{homeier1996} which are products of the Clebsch-Gordan
  coefficients or Wigner 3-j symbols. The Gaunt coefficients have a closed form
  expression in terms of $j,j'$, and $j''$ (usually expressed in terms of
  $l,l',l'',m,m',m''$) that is implemented in the Sympy library \cite{sympy}.

  To calculate the kernel for general polarized light microscopes we will need to multiply functions in the following form
  \begin{align}
  f(\mh{p}, \mh{s}) = \sum_{n=0}^\infty\sum_{j=0}^\infty c_{n,j} z_n(\mh{p})y_j(\mh{s}), \qquad f'(\mh{p}, \mh{s}) = \sum_{n'=0}^\infty\sum_{j'=0}^\infty c_{n',j'}' z_n(\mh{p})y_j(\mh{s}).
\end{align}
The product will be in the form
\begin{align}
  f(\mh{p}, \mh{s})f'(\mh{p}, \mh{s}) = \sum_{n=0}^\infty\sum_{j=0}^\infty c_{n'',j''}'' z_{n''}(\mh{p})y_{j''}(\mh{s}),
\end{align}
where
\begin{align}
  {c''}^{n'', j''} = P_{n,n'}^{n''} G_{j,j'}^{j''} c_{n,j} c_{n', j'}'.\label{eq:final}
\end{align}
Equation \ref{eq:final} is the main result of this section. It shows that we can
precalculate the triple integrals of the circular and spherical harmonics and
use the results to efficiently find the coefficients of the product of two
arbitrary kernels. We can think of Eq. \ref{eq:final} as a bilinear map that
acts within the vector space of harmonic function coefficients. The bilinear map
takes two elements of the vector space and maps them to another element of the
vector space by a rank-6 tensor product. A lower dimensional example of a
bilinear map is the cross product which takes two vectors in three-dimensional
Euclidean space and maps them to another vector by a rank-3 tensor product.
\end{document}
